---
title: "Integrating pig farming into cocoa productionin the south region of Cameroon"
author: "Justice bill Tcheutchoua Kengoum (50275546)"
date: "2025-07-18"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Introduction**
Cameroon is a notable player in Africa’s agricultural sector, with cocoa and pig production representing vital components of its rural economy. Cocoa is the country’s leading export crop, contributing substantially to national revenue and providing livelihoods for thousands of smallholder farmers (FAO, 2023).It is a male dominant with limited land ownership and access to extension services. It is profitable with pooled Net Farm Income (NFI) of 79485 FCFA/ha and farms within 10 to 30years give highest yields. In 2022, Cameroon produced approximately 250,000 to 300,000 tonnes of cocoa, positioning it among the top cocoa producers in Africa (ITC, 2023). Despite its economic significance, the cocoa sector faces challenges such as aging trees, climate change, disease outbreaks like cocoa pod borer, and climate variability impacting yields (Nkouamé et al., 2021). Efforts to enhance productivity include adopting improved plant varieties and processing technologies. 
Pig production, predominantly based on smallholder farms, also plays a crucial role in Cameroon’s livestock industry, fulfilling domestic demand and supplementing rural livelihoods (FAO, 2023). The sector faces constraints such as limited access to quality inputs, veterinary services, and disease management, notably African swine fever outbreaks (Meli et al., 2022). Nevertheless, increasing interest in breed improvement and intensive husbandry practices offers potential for growth.Records from the ministry of livestock in 2022 and 2023 declare 46000 and 49400 tons of pork production. The Annual meat consumption is 13.3kg/inhabitants/year and pork contribution to this consumption is 2.02kg/inhabitants/year. Standard recommendation by FAO is 42kg/inhabitants/year, with pork contributing 5kg/inhabitants/year. This shows a deficit of about 28.7kg of meat and 11.1kg of pork/inhabitants/year, hence production is not sufficient to meet the demand. The most common production system carried out here is traditional and is mostly done by male, aged between 20 to 40 years. Most of them are involved in other agricultural activities and 19% are formally trained. These farmers are mostly involved in both fattening and breeding. 72% of these farmers rely on both household remains and concentrate for feeding their animals.
Integrating pig and cocoa farming in Cameroon offers significant socio-economic and political benefits. This agro-ecological approach enhances income diversification for smallholder farmers, reducing reliance on monoculture and increasing resilience (Momo et al., 2020). Pigs can efficiently recycle cocoa pod husks as feed, promoting sustainable waste management (Ngatcha et al., 2019). Politically, this integration aligns with Cameroon’s development goals to boost rural livelihoods and food security (MINADER, 2022). Economically, it creates employment opportunities and supports local markets. Socially, it fosters community cooperation and gender inclusion, vital for rural development. Overall, this integrated model is crucial for sustainable agricultural growth in Cameroon.

## **Obectives**
The aim of this project is to determine whether integrating pig farming into cocoa production will generate supplementary benefits to encourage existing cocoa farmers to venture into the new production system. This will be achieved by doing the following;determine the cost of production, total revenue, total profit and Net Present Value (NPV) of the 2 main production systems (cocoa production and integrated pig farming in to cocoa production)

## **Methodology**
Decision analysis is a systematic approach for making decisions under uncertainty, incorporating mathematical and analytical techniques to evaluate alternative courses of action and their potential outcomes. It is a quantitative approach that aims to support decision-making processes by explicitly considering uncertainties, trade-offs, and potential consequences. in this research work, we have to;
- Identify the problem; Here, we shall Clearly defining the decision problems and identify the objectives, constraints, and all stakeholders involved. 
- Collect sufficent data in pig and cocoa production from both experts and farmers.
- Develop a model to represent the relationship between all the variables involved.
- Attibute risks to the model; here we shall consider the effects (negative) affecting production and to what extend the results (output or profit) changes when considering it in our mathematical model.
- Analyse and evaluate the results obtained; this will be done in Rstudio (aminly statistical analysis)
- Recommend, sensitize, advertize the results obtained for implementation; make available the results of the research work to farmers and stakeholders.
This work was mainly done with the use of R software version 4.3.2. It has several packages that will be used to help in the analysis, one of which is the decisionSupport R package. It provides a comprehensive set of functions and
tools to support decision analysis using the R programming language. It offers a flexible and customizable platform for implementing decision support methodologies, enabling users to perform sophisticated decision analysis tasks
efficiently.

# decision support package used
```{r package-management, echo=TRUE}
# load libraries:
library(decisionSupport)
library(dplyr)
library(ggplot2)
library(readr)
library(knitr)

#Automatically write R package citation entries to a .bib file
#knitr::write_bib(c(.packages(),
#                   'decisionSupport',
#                  'dplyr',
#                   'ggplot2'),'project_packages.bib')
```

# **Decision Criteria**
The decision makers in this research work are farmer cocoa farmers. They are the ones having land and cocoa farms on which the pigs will be integrrated. since the majority of farmers are smallholders with low education standard, they always prefer to see results before adopttion and implementation, we are therefore looking for cocoa farmers with farm size between 2 to 5 hectare. These farm must be in production for more than 10 years and not less than 30 since cocoa farms are in maximum production at this stage and it is possible to have clear results working with them.
- baseline: cocoa farmers
- Decision maker:Smallholder cocoa Farmers
- criteria : cocoa farmers with farm size between 2 to 5ha and are in the activity for over 10 to 15 years of production.
- Time span: 5 years

# **Conceptual model**

https://files.slack.com/files-pri/T01TDL4J7KL-F095BCJRXQW/conceptual_model_i.drawio__5_.png

# **Data collection (input data)**
```{r load_input_data}

## Load required packages
library(decisionSupport)
library(ggplot2)

library(readr)
getwd()

#Read the unified input table
input_data <- read.csv("C:/Users/willy/Desktop/COCAPIG/fsdfg/PC.csv")
input_data
# Convert to estimate object and draw one set of inputs into the environment
estimates <- as.estimate(input_data)
make_variables <- function(est, n = 1) {
  x <- random(rho = est, n = n)
  for (nm in colnames(x)) {
    assign(nm, as.numeric(x[1, nm]), envir = .GlobalEnv)
  }
}

make_variables(estimate_read_csv(paste("C:/Users/willy/Desktop/COCAPIG/fsdfg/PC.csv")))
```
# **Data analysis**
```{r model_function}
model_function <- function(x, input_data) {
  ### 2.1  cocoa cost in monoculture
  cocoa_bean_co1 <- vv(cocoa_bean_co, var_CV = var_CV, n = n_years)
  tran_cost_co1 <- vv(tran_cost_co, var_CV = var_CV, n = n_years)
  fer_cost_co1 <- vv(fertiliser_cost_co, var_CV = var_CV, n = n_years)
  lab_cost_co1 <- vv(labor_co, var_CV = var_CV, n = n_years)
  farm_man_co1 <- vv(farm_man_co, var_CV = var_CV, n = n_years)
  
  cocoa_co_tc <- cocoa_bean_co1+tran_cost_co1+fer_cost_co1+
              farm_man_co1+lab_cost_co1
  
    ### 2.1.1 risk  effects in cocoa production
  dry_year <- chance_event(
    chance       = dry_year_co_d,
    value_if     = Coco_Yield_co * (1 - dry_year_co_e),
    value_if_not = Coco_Yield_co,
    n            = n_years )
  
  heavt_rain <- chance_event(
    chance       = heavy_rain_co_d,
    value_if     = Coco_Yield_co * (1 - heavy_rain_co_e),
    value_if_not = Coco_Yield_co,
    n            = n_years )
  
  draugh_year <- chance_event(
    chance       = draught_co_d,
    value_if     = Coco_Yield_co * (1 - draught_co_e),
    value_if_not = Coco_Yield_co,
    n            = n_years )
  
  pest_year <- chance_event(
    chance       = pest_co_d,
    value_if     = Coco_Yield_co * (1 - pest_co_e),
    value_if_not = Coco_Yield_co,
    n            = n_years )
  
  cocoa_total_risk <- dry_year + heavt_rain + draugh_year + pest_year
  
  
  ### 2.1.2 cocoa revenue after loss
  cocoa_revenue_with_risks <- vv(
    var_mean = (Coco_Yield_co * Coco_Price_co) - (cocoa_total_risk),
    var_CV   = var_CV,
    n        = n_years )
  
  cocoa_revenue_No_risk <-  vv(
    var_mean = (Coco_Yield_co * Coco_Price_co),
    var_CV   = var_CV,
    n        = n_years )
 
  ### cocoa Cashflow & NPV
  cocoa_Cashflow <- cocoa_revenue_with_risks - cocoa_co_tc
  CumCashflow_cocoa <- cumsum(cocoa_Cashflow)
  NPV_cocoa         <- discount(cocoa_Cashflow,
                                      discount_rate = discount_rate,
                                      calculate_NPV = TRUE)
  
  # 2.2 integrated production
  
  ### 2.2.1 cocoa and pig initial cost
  farm_inst_cp1 <- vv(farm_inst_cp, var_CV = var_CV, n = n_years)
  mat_eq_cp1 <- vv(mat_eq_cp, var_CV = var_CV, n = n_years)
  water_prov_cp1 <- vv(water_prov_cp, var_CV = var_CV, n = n_years)
 
  cocoa_cp_tci <- farm_inst_cp1 + mat_eq_cp1 + water_prov_cp1
  
  ### 2.2.2 cocoa and pig Operational costs 
  biop_cp1 <- vv(biop_cp, var_CV = var_CV, n = n_years)
  pig_feed_cp1 <- vv(pig_feed_cp, var_CV = var_CV, n = n_years)
  veter_cost_cp1 <- vv(veter_cost_cp, var_CV = var_CV, n = n_years)
  shared_lab_cp1 <- vv(shared_lab_cp, var_CV = var_CV, n = n_years)
  tran_cost_cp1 <- vv(tran_cost_cp, var_CV = var_CV, n = n_years) 
  cocoa_bean_cp1 <- vv(cocoa_bean_cp, var_CV = var_CV, n = n_years)
  husk_cp1 <- vv(husk_cp, var_CV = var_CV, n = n_years)
  biof_cp1 <- vv(biof_cp, var_CV = var_CV, n = n_years)
  
  cocoa_cp_tco <- biop_cp1 + pig_feed_cp1 + veter_cost_cp1 + shared_lab_cp1 +
                   tran_cost_cp1 + cocoa_bean_cp1 + husk_cp1 + biof_cp1 
  
  cocoa_and_pig_total_cost_in_year_1 <- cocoa_cp_tci + cocoa_cp_tco

  ##  2.2.3 Risk adjustment in integrated cocoa and pig
  
  theft_year <- chance_event(
    chance       = theft_cp_d,
    value_if     = Pig_Yield_cp * (1 - theft_cp_e),
    value_if_not = Pig_Yield_cp,
    n            = n_years )
  
  pest_year <- chance_event(
    chance      = pest_cp_d,
    value_if    = Pig_Yield_cp * (1 - pest_cp_e),
    n           = n_years)
  
  dry_year <- chance_event(
    chance      = dry_year_cp_d,
    value_if    = Coco_Yield_cp * (1- dry_year_cp_e) ,
    n           = n_years)
  
  heavy_rain_year <-  chance_event(
    chance      = heavy_rain_cp_d,
    value_if    = Coco_Yield_cp * (1- heavy_rain_cp_e) ,
    n           = n_years)
  
  draugh_year <- chance_event(
    chance      = dry_year_cp_d,
    value_if    = Coco_Yield_cp * (1- dry_year_cp_e) ,
    n           = n_years)
  
  Integrated_cp_total_risks <-  theft_year + pest_year + dry_year + heavy_rain_year + draugh_year
  
    ### 2.3integrated cocoa and pig revenue
  integrated_revenue_No_risks <- vv( var_mean = (Pig_Yield_cp * pig_Price_cp) +
                     (Coco_Yield_cp * Coco_Price_cp),
                     var_CV   = var_CV,
                     n        = n_years )
  
  integrated_revenue_with_risks <- vv( var_mean = ((Pig_Yield_cp * pig_Price_cp) +
                                  (Coco_Yield_cp * Coco_Price_cp)) - Integrated_cp_total_risks,
                                  var_CV   = var_CV,
                                  n        = n_years )
  
    ### Integrated pig and cocoa Cashflow & NPV
  Cashflow_integrated_cocoa_and_pig    <- vv(integrated_revenue_with_risks - (cocoa_cp_tci), 
                                         var_CV   = var_CV,
                                          n        = n_years ) 
    CumCashflow_integrated_cocoa_and_pig <- cumsum(Cashflow_integrated_cocoa_and_pig)
  NPV_Integrated_cocoa_and_pig         <- discount(Cashflow_integrated_cocoa_and_pig,
                                        discount_rate = discount_rate,
                                        calculate_NPV = TRUE)
  
  ### Decision difference: Profit, cashflow, cumulative
  Cashflow_Decision    <- Cashflow_integrated_cocoa_and_pig - cocoa_Cashflow
  CumCashflow_Decision <- cumsum(Cashflow_Decision)
  NPV_Decision         <- NPV_Integrated_cocoa_and_pig - NPV_cocoa
  
  
  ## Return all outputs
  return(list(
    NPV_cocoa        = NPV_cocoa,
    NPV_Integrated_cocoa_and_pig      = NPV_Integrated_cocoa_and_pig,
    NPV_Decision           = NPV_Decision,
    cocoa_Cashflow   = cocoa_Cashflow,
    Cashflow_integrated_cocoa_and_pig = Cashflow_integrated_cocoa_and_pig,
    Cashflow_Decision      = Cashflow_Decision,
    CumCashflow_cocoa   = CumCashflow_cocoa,
    CumCashflow_integrated_cocoa_and_pig = CumCashflow_integrated_cocoa_and_pig,
    CumCashflow_Decision      = CumCashflow_Decision
  ))
}
```

# **Running monto carlo simulation(MC simulation)**
The model function was run 10000 times to estimate risks, cost and benefits from the decision model function
```{r MC simulation}
# Run a single Monte Carlo simulation for both scenarios
combined_simulation <- mcSimulation(
  estimate = estimate_read_csv("C:/Users/willy/Desktop/COCAPIG/fsdfg/PC.csv"),
  model_function    = model_function,
  numberOfModelRuns = 10000,
  functionSyntax    = "plainNames" )

combined_simulation
```

# **Results visualisation**
## **Distribution of NPVs**
## **NPV in cocoa production system**
```{r NPV in cocoa production system}
plot_distributions(
  mcSimulation_object = combined_simulation,
  vars                = "NPV_cocoa",
  method              = "boxplot_density", base_size = 7,
  x_axis_name = "Outcome distribution (Fcfa/ha)")

```

## **NPV in integrated cocoa and pig production system**
```{r NPV in cocoa and pig production system}
plot_distributions(
  mcSimulation_object = combined_simulation,
  vars                = "NPV_Integrated_cocoa_and_pig",
  method              = "boxplot_density", base_size = 7,
  x_axis_name = "Outcome distribution (Fcfa/ha)")

```
## **NPV in cocoa and integrated pig and cocoa system**
```{r NPV for cocoa and integrated pig and cocoa production system}
plot_distributions(
  mcSimulation_object = combined_simulation,
  vars                = c("NPV_cocoa","NPV_Integrated_cocoa_and_pig"),
  method              = "smooth_simple_overlay", base_size = 7,
  x_axis_name = "Outcome distribution (Fcfa/ha)")
```

## **NPV for decision analysis**
```{r NPV and cashfloe for cocoa}
# Compound figure of NPV and cashflow for cocoa
compound_figure(
  # mcSimulation_object = combined_simulation,
  model = model_function,
  input_table = input_data,
  decision_var_name = "NPV_cocoa",
  cashflow_var_name = "cocoa_Cashflow",
  model_runs = 1000, 
  distribution_method = 'smooth_simple_overlay')
```
```{r combine inputs and both NPVS into one data frame}
# Combine inputs and both NPVs into one data frame
df_evpi <- data.frame(
           combined_simulation$x,
           NPV_cocoa   = combined_simulation$y[, "NPV_cocoa"],
           NPV_Integrated_cocoa_and_pig = combined_simulation$y[, "NPV_Integrated_cocoa_and_pig"] )



# Calculate EVPI for all inputs, starting at NPV_cocoa
EVPI_results <- multi_EVPI(
                mc            = df_evpi,
                first_out_var = "NPV_cocoa" )

# Inspect the EVPI table
print(EVPI_results)

# Plot EVPI for both scenarios side by side
plot_evpi(
         EVPIresults   = EVPI_results,
         decision_vars = c("NPV_cocoa", "NPV_Integrated_cocoa_and_pig"))
```


## **Annual and Cumulative Cashflows**

### **Annual Cumulative CashFlow**
```{r annual cummulative cashflow}
# Plot the annual profit (“cashflow”) time series for both scenarios
plot_cashflow(
  mcSimulation_object = combined_simulation,
  cashflow_var_name   = c("cocoa_Cashflow", "Cashflow_integrated_cocoa_and_pig"),
  x_axis_name         = "Year",
  y_axis_name         = "Annual Profit (Fcfa/ha)",
  color_25_75         = "blue",
  color_5_95          = "lightblue",
  color_median        = "darkblue",
  facet_labels        = c("cocoa", "integrated_cocoa_and_pig"))

plot_cashflow(
  mcSimulation_object = combined_simulation,
  cashflow_var_name   = c("CumCashflow_cocoa", "CumCashflow_integrated_cocoa_and_pig"),
  x_axis_name         = "Year",
  y_axis_name         = "Annual Profit (Fcfa/ha)",
  color_25_75         = "blue",
  color_5_95          = "lightblue",
  color_median        = "darkblue",
  facet_labels        = c("cocoa", "integrated_cocoa_and_pig"))
```

## **EVPI Analysis**
```{r evpi-analysis, echo=TRUE, message=FALSE, warning=FALSE}
# Calculate EVPI for all inputs, starting at NPV_cocoa
EVPI_results <- multi_EVPI(
                mc            = df_evpi,
                first_out_var = "NPV_cocoa" )

# Inspect the EVPI table
print(EVPI_results)

# Plot EVPI for both scenarios side by side
plot_evpi(
         EVPIresults   = EVPI_results,
         decision_vars = c("NPV_cocoa", "NPV_Integrated_cocoa_and_pig"))
```


### **Additional Result Graphs**

```{r Additional Result Graphs}
#### Additional Result Graphs ####

# Side-by-side boxplots of NPV
NPV_df <- data.frame(
          cocoa   = combined_simulation$y[, "NPV_cocoa"],
          integrated_pig_and_cocoa = combined_simulation$y[, "NPV_Integrated_cocoa_and_pig"])

NPV_long <- pivot_longer(
            NPV_df,
            cols      = everything(),
            names_to  = "Scenario",
            values_to = "NPV")

ggplot(NPV_long, aes(x = Scenario, y = NPV, fill = Scenario)) +
       geom_boxplot(alpha = 0.6) +
       theme_minimal(base_size = 14) +
       labs(
       title = "NPV Distribution: cocoa vs. integrated_pig_and_cocoa",
       x     = NULL,
       y     = "Net Present Value (Fcfa/ha)"
       ) +
       scale_fill_manual(values = c("grey40", "forestgreen")) +
       theme(legend.position = "none")


# Grab columns
cocoa_cols <- grep("^CumCashflow_cocoa",   names(combined_simulation$y),
                  value = TRUE)
integrated_pig_and_cocoa_cols  <- grep("^CumCashflow_integrated_cocoa_and_pig", names(combined_simulation$y), 
                  value = TRUE)
dec_cols  <- grep("^CumCashflow_Decision",      names(combined_simulation$y), 
                  value = TRUE)

# Sanity check
if (!(length(cocoa_cols)==n_years && length(integrated_pig_and_cocoa_cols)==n_years && 
      length(dec_cols)==n_years)) {
      stop(sprintf("Expect %d years, got %d mono, %d int, %d dec",
      n_years, length(mono_cols), length(integrated_pig_and_cocoa_cols), length(dec_cols)))}

# Compute medians per year
cocoa_med_cum <- apply(combined_simulation$y[, cocoa_cols], 2, median)
integrated_pig_and_cocoa_med_cum  <- apply(combined_simulation$y[, integrated_pig_and_cocoa_cols],  2, median)
dec_med_cum  <- apply(combined_simulation$y[, dec_cols],  2, median)


# Build long data.frame for just cocoa & integrated_cocoa_and_pig
year_seq <- seq_len(n_years)
cum_df <- data.frame(
          Year       = rep(year_seq, times = 2),
          Cumulative = c(cocoa_med_cum, integrated_pig_and_cocoa_med_cum),
          Scenario   = rep(c("cocoa","integrated_pig_and_cocoa"), each = n_years))

# Plot it
ggplot(cum_df, aes(x = Year, y = Cumulative, color = Scenario)) +
      geom_line(size = 1.2) +
      geom_point(size = 2) +
      theme_minimal(base_size = 14) +
      labs(
      title = "Median Cumulative Profit Over Time",
      x     = "Year",
      y     = "Cumulative Profit (USD/ha)"
      ) +
      scale_color_manual(values = c("grey40","forestgreen"))

```

# Discussion
The findings of this research work showed that the cocoa annaul profit varies from 0 to 15 millions with an average of 3 to 8 millions. The minimum cummulative annual profit of 200000 to 18000000 and maxumum of 1 to 75 millions in 5 years.the NPV varies from 12.5 to 37.5 millions.
on the other hand, the integrated cocoa and pig annaul profit varies from 0 to 18 millions with an average of 3.6 to 11 millions. The minimum cummulative annual profit of 3 millions to 5 millions and maxumum of 10 to 37.55 millions in 5 years.the NPV varies from 16 to 42 millions.
These results clearly shows a greater increase in profit in integrated production systems than cocoa production systems.
# Conclusion
integrating pig farming into cocoa production gives better profit and should therefore be adopted by farmers in the south region of Cameroon.

# References







